# **title: "Reproducible Research: Peer Assessment 1"**
output: 

html_document:

keep_md: true

**Loading and preprocessing the data (Code for reading in the dataset and/or processing the data)**

download file from web
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")
```
unzip data and read 
```{r}
unzip("activity.zip")
stepdata <- read.csv("activity.csv", header = TRUE)
head(stepdata)
```
**What is mean total number of steps taken per day?**

Histogram of the total number of steps taken each day
```{r}
library(magrittr)
library(dplyr)
databydate <- stepdata %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps= sum(steps)) %>%na.omit()
hist(databydate$tsteps, col = "pink" , xlab = "Number of steps per day",main="Total number of steps taken each day", breaks = 20)
```

**Mean and median number of steps taken each day**
```{r}
mean(databydate$tsteps)
print(mean(databydate$tsteps))
median(databydate$tsteps)
print(median(databydate$tsteps))
```
**What is the average daily activity pattern?**

Time series plot of the average number of steps taken
```{r}
databyinterval <- stepdata%>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(tsteps= mean(steps))
plot(databyinterval, type = "l", 
     main = "Average daily activity pattern", 
     ylab = "Avarage number of steps taken", 
     xlab = "5-min intervals")
```

The 5-minute interval that, on average, contains the maximum number of steps
```{r}
interval_row <- which.max(databyinterval$tsteps)
max_interval <- databyinterval[interval_row,1:2]
print(max_interval)
```
**Imputing missing values (Code to describe and show a strategy for imputing missing data)**

Calculating and reporting the total number of missing values in the data set
```{r}
missingVals <- sum(is.na(stepdata))
print(missingVals)
```
Devising a strategy for filling in all of the missing values in the data set
```{r}
library(magrittr)
library(dplyr)

replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
print(meandata)
```
Histogram of the total number of steps taken each day after missing values are imputed
```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)
print(FullSummedDataByDay,15)
summary(FullSummedDataByDay)
print(summary(FullSummedDataByDay))
hist(FullSummedDataByDay$totalsteps, col = "violet", xlab = "Number of Steps per day", ylab = "Frequency", main = "Total number of steps taken each day", breaks = 20)
```

Calculating and reporting the mean and median : total number of steps taken per day
```{r}
oldmean <- mean(databydate$tsteps, na.rm = TRUE)
newmean <- mean(FullSummedDataByDay$totalsteps)
```
Old mean and New mean
```{r}
print(oldmean)
print(newmean)
oldmedian <- median(databydate$tsteps, na.rm = TRUE)
newmedian <- median(FullSummedDataByDay$totalsteps)
```
Old median and New median
```{r}
print(oldmedian)
print(newmedian)
```
**Are there differences in activity patterns between weekdays and weekends?**
```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )
library(ggplot2)
meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")

plot(ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + 
  geom_line() + 
  facet_grid(weekend ~.) +
  xlab("5-minute intervals") + 
  ylab("Avarage number of steps taken") +
  ggtitle("Weekdays and weekends activity patterns"))
```